Query Optimization for Dynamic Imputation

### 2017-03-01 paper/main.tex
Move planing times results to commands
paper: null made macro, since had null and NULL
paper: some minor fixes and todos added
paper: add equal contribution mark

\author{
  Jos\'e Cambronero \\
  \texttt{jcamsan@csail.mit.edu}
  \and
  John K. Feser \\
  \texttt{feser@csail.mit.edu}
  \and
  Micah J. Smith \\
  \texttt{micahs@mit.edu}
  \and
  Samuel Madden \\
  \texttt{madden@csail.mit.edu}}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\todo}[1]{{\color{red}\textbf{TODO:} #1}}
\newcommand{\todobox}[2]{{\color{red}\fbox{\parbox{\columnwidth}{\textbf{TODO:} #1\\#2}}}}

% results that can be used in any section
\newcommand{\demorows}{10175}
\newcommand{\labexrows}{9813}
\newcommand{\acsbaseresultminutes}{355 minutes}
\newcommand{\acsbaseresulthours}{just under 6 hours}
\newcommand{\acsimputedbzeroresult}{4 seconds}
\newcommand{\acsimputedboneresult}{1 second}
\newcommand{\lowxalphazero}{10}
\newcommand{\lowsmapealphazero}{0}
\newcommand{\highxalphazeroexacs}{200}
\newcommand{\highxalphaoneexacs}{1000}
\newcommand{\highsmapealphaoneexacs}{24}
\newcommand{\highxalphaone}{16000}

@@

\begin{abstract}
Missing values abound in data analysis and often present a serious challenge for usability of data. Users
are forced to pick between removing tuples with missing values or creating a clean duplicate of their data
by  applying a potentially-expensive imputation strategy. Our system, \ProjectName{}, incorporates imputation into a cost-based query optimizer, 
performing imputation on the fly
for each query. This allows users to immediately explore
their data, while the system picks the optimal placement of imputation operations. This placement is based
on the specific query issued and a user-provided parameter that trades off data
quality and runtime performance. We incorporate the imputation operators into the relational algebra, allowing
for traditional optimizations to take place. Our experiments evaluate this approach on two real-world survey-based datasets
from the \textit{Center for Disease Control and Prevention} and \textit{freeCodeCamp}. Our
results show that the optimizer generates query plans that execute between \lowxalphazero{}
and \highxalphazeroexacs{} times as fast as imputing at the base tables and then executing
queries, making exploratory data analysis using our system feasible.  Furthermore, we show
that the query results from on-the-fly imputation have
Symmetric-Mean-Absolute-Percentage-Error between \lowsmapealphazero{} and
\highsmapealphaoneexacs{} percent when compared to standard full-table imputation
approach.

>>>>

\author{
  Jos\'e Cambronero\thanks{Author contributed equally to this paper.} \\
  \texttt{jcamsan@csail.mit.edu}
  \and
  John K. Feser\footnotemark[1] \\
  \texttt{feser@csail.mit.edu}
  \and
  Micah J. Smith\footnotemark[1] \\
  \texttt{micahs@mit.edu}
  \and
  Samuel Madden \\
  \texttt{madden@csail.mit.edu}}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\todo}[1]{{\color{red}\textbf{TODO:} #1}}
\newcommand{\todobox}[2]{{\color{red}\fbox{\parbox{\columnwidth}{\textbf{TODO:} \textit{#1}\\#2}}}}

% results that can be used in any section
\newcommand{\demorows}{10175}
\newcommand{\labexrows}{9813}
\newcommand{\acsbaseresultminutes}{355 minutes}
\newcommand{\acsbaseresulthours}{just under 6 hours}
\newcommand{\acsimputedbzeroresult}{4 seconds}
\newcommand{\acsimputedboneresult}{1 second}
\newcommand{\lowxalphazero}{10}
\newcommand{\lowsmapealphazero}{0}
\newcommand{\highxalphazeroexacs}{200}
\newcommand{\highxalphaoneexacs}{1000}
\newcommand{\highsmapealphaoneexacs}{24}
\newcommand{\highxalphaone}{16000}
\newcommand{\nullv}{NULL}
\newcommand{\planningruntimepercent}{8.5}
\newcommand{\planningmaxtime}{40}

@@

\begin{abstract}
Missing values abound in data analysis and often present a serious challenge for usability of data. Users
are forced to pick between removing tuples with missing values or creating a clean duplicate of their data
by applying a potentially-expensive imputation strategy. Our system, \ProjectName{}, incorporates imputation into a cost-based query optimizer, 
performing any operations needed on the fly
for each query. This allows users to immediately explore
their data, while the system picks the optimal placement of imputation operations. This placement is based
on the specific query issued and a user-provided parameter that trades off data
quality and runtime performance. We incorporate the imputation operators into the relational algebra, allowing
for traditional optimizations to take place. Our experiments evaluate this approach on two real-world survey-based datasets
from the \textit{Center for Disease Control and Prevention} and \textit{freeCodeCamp}. Our
results show that the optimizer generates query plans that execute between \lowxalphazero{}
and \highxalphazeroexacs{} times as fast as imputing at the base tables and then executing
queries, making exploratory data analysis using our system feasible.  Furthermore, we show
that the query results from on-the-fly imputation have
Symmetric-Mean-Absolute-Percentage-Error between \lowsmapealphazero{} and
\highsmapealphaoneexacs{} percent when compared to standard full-table imputation
approach.
